### 2020-09-08 proceedings/tex/sections/appendix.tex

\appendix

\begin{lstlisting}[caption={Program using Causal.jl for the simulation of the system in Figure \ref{fig: coupled model}. Plots.jl is used to plot the simulation data while FFTW.jl is used to compute the FFT of the simulation data \cite{plots, fftw}.}, label={lst: coupled codes}, language=Julia]
using Causal, Plots, FFTW

# Define ODE system type under AbstractODESystem
@def_ode_system mutable struct ODESys{RH, RO, IP, OP} <: AbstractODESystem
    righthandside::RH = function fode(dx, x, u, t, a=35., b=3., c=28.)
        dx[1] = a * (x[2] - x[1])
        dx[2] = (c - a) * x[1] + c * x[2] - x[1] * x[3]
        dx[3] = x[1] * x[2] - b * x[3]
    end
    readout::RO = (x, u, t) -> x 
    state::Vector{Float64} = rand(3)
    input::IP = nothing
    output::OP = Outport(3)
end

# Define SDE system type under AbstractSDESystem
@def_sde_system mutable struct SDESys{DR, DF, RO, IP, OP} <: AbstractSDESystem
    drift::DR = function fsde(dx, x, u, t, a=35., b=3., c=28.)
        dx[1] = a * (x[2] - x[1]) + u[1](t)
        dx[2] = (c - a) * x[1] + c * x[2] - x[1] * x[3] + u[2](t)
        dx[3] = x[1] * x[2] - b * x[3] + u[3](t)
    end
    diffusion::DF = (dx, x, u, t, η=10.) -> (dx .= η) 
    readout::RO = (x, u, t) -> x
    state::Vector{Float64} = rand(3)
    input::IP = Inport(3)
    output::OP = Outport(3)
end

# Define Gain system type under AbstractStaticSystem
@def_static_system struct GainSys{RO,IP,OP} <: AbstractStaticSystem 
    readout::RO = (u,t, ϵ=0.1) -> ϵ * u 
    input::IP = Inport(3)
    output::OP = Outport(3)
end

# Define plugin type under AbstractPlugin
@def_plugin struct FFTPlug{PR} <: AbstractPlugin 
    process::PR = x -> fft(x, 2)    # Computes FFT of the data `x`
end

# Construct model
@defmodel model begin 
    @nodes begin 
        odeds = ODESys()    # Add odeds component  
        gain = GainSys()    # Add gain  component   
        sdeds = SDESys()    # Add sde components
        writer1 = Writer(input=Inport(3))                   # Add writer without plugin
        writer2 = Writer(input=Inport(3))                   # Add writer without plugin
        writer3 = Writer(input=Inport(3), plugin=FFTPlug()) # Add writer with `FFTPlug` plugin
    end
    @branches begin 
        odeds => gain       # Add connection from `odeds` to `gain`.
        gain => sdeds       # Add connection from `gain` to `sdeds`.
        odeds => writer1    # Add connection from `odeds` to `writer1`.
        sdeds => writer2    # Add connection from `sdeds` to `writer2`.
        odeds => writer3    # Add connection from `odeds` to `writer3`.
    end 
end

# Simulate model 
sim = simulate!(model, 0, 0.001, 100) # Simulate from 0 to 100 seconds with a step size of 0.001 seconds

# Plot simulation data 
t, x1 = read(getnode(model, :writer1).component)
t, x2 = read(getnode(model, :writer2).component)
x3 = read(getnode(model, :writer3).component, flatten=false)
plot(x1[:, 1], x1[:, 2], subplot=1, label="")
plot(x2[:, 1], x2[:, 2], subplot=2, label="")   
println(x3)  
\end{lstlisting}

\begin{lstlisting}[caption={Program using Causal.jl for the simulation of the system in Figure \ref{fig: network graph}. Plots.jl is used to plot the simulation data while LightGraphs.jl is used to construct the connectiviy matrix of the network topology \cite{plots, lightgraphs}}, label={lst: network codes}, language=Julia]
using Causal, Plots
using LightGraphs: watts_strogatz, laplacian_matrix

# Define DynamicSystem type under AbstractODESystem
@def_ode_system mutable struct DynamicSystem{RH, RO, IP, OP} <: AbstractODESystem
    righthandside::RH = function f(dx, x, u, t, σ=10., β=8/3, ρ=28)
        dx[1] = σ * (x[2] - x[1]) + u[1](t)
        dx[2] = x[1] * (ρ - x[3]) - x[2] + u[2](t)
        dx[3] = x[1] * x[2] - β * x[3] + u[3](t)
    end
    readout::RO = (x, u, t) -> x
    state::Vector{Float64} = rand(3)
    input::IP = Inport(3)
    output::OP = Outport(3) 
end

# Define Coupler type under AbstractStaticSystem
@def_static_system struct CouplerSystem{ET, CT, IP, OP, RO} <: AbstractStaticSystem
    E::ET
    P::CT
    input::IP = Inport(size(E, 1) * size(P, 1))
    output::OP = Outport(size(E, 1) * size(P, 1))
    readout::RO =  (u, t, E=E, P=P) ->  kron(E, P) * u
end

# Construct network parameters 
n = 50      # Number of nodes 
d = 3       # Dimension of each node
ϵ = 20.     # Coupling strength 
components = [DynamicSystem() for i in 1 : n]
E = ϵ * collect(-laplacian_matrix(watts_strogatz(n, 10, 0.5)))  # Connectivity matrix of network topology
P = [1 0 0; 0 0 0; 0 0 0]   # Systems are coupled by their first state variable.

# Construct model 
model = Model()

# Add components
for (i, component) in enumerate(components)
    addnode!(model, component, label=Symbol("node$i"))
end
addnode!(model, CouplerSystem(E=E, P=P), label=Symbol("coupler"))
addnode!(model, Writer(input=Inport(n * d)), label=Symbol("writer"))

# Add connections
m = n + 1
for (j, k) in zip(1:n, map(i -> i : i + d - 1, 1 : d : n * d))
    addbranch!(model, j => m, 1 : d => k)
    addbranch!(model, m => j, k => 1 : d)
end
p = n + 2
for (j, k) in zip(1:n, map(i -> i : i + d - 1, 1 : d : n * d))
    addbranch!(model, j => p, 1 : d => k)
end

# Simulate model 
sim = simulate!(model, 0, 0.001, 100.)

# Read and plot the simulation data 
t, x = read(getnode(model, :writer).component)

# Compute mean square error
mse = sqrt.(sum(diff(x[:, 1 : d : end],dims=2).^2/n, dims=2))

# Plot results
k = 3000 
plot(t, x[:, 1] , label="")
plot(x[:, 1], x[:,2], label="")
plot(t[1:k], mse[1:k], label="")
\end{lstlisting}