### 2020-09-08 proceedings/tex/main.tex
Renamed file name [skip ci]

\documentclass{juliacon}
\setcounter{page}{1}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{mathtools}
\usepackage{cite}

\title{Causal.jl:\\A Modeling and Simulation Framework for Causal Models}

\author[1]{Zekeriya Sarı}
\author[1]{Serkan Günel}
\affil[1]{Department of Electrical and Electronics Engineering, Dokuz Eylül University, İzmir, Turkey}

\keywords{System Modeling, Causal Modeling, System Simulation, Dynamical Systems, Julia Programming Language}

\hypersetup{
pdftitle = {Causal.jl: A Modeling and Simulation Framework for Causal Models},
pdfsubject = {JuliaCon 2020 Proceedings},
pdfauthor = {1st author, 2nd author, 3rd author},
pdfkeywords = {System Simulation, System Modeling, Dynamical Systems, Julia Programming Language},
}

\hypersetup{
  colorlinks   = true,  % Colours links instead of ugly boxes
  urlcolor     = black,  % Colour for external hyperlinks
  linkcolor    = black,   % Colour of internal links
  citecolor   = black     % Colour of citations
}

\begin{document}
    \maketitle
    \input{sections/abstract.tex}
    \input{sections/introduction.tex}
    \input{sections/modelling_and_simulation.tex}
    \input{sections/illustrative_examples.tex}
    \input{sections/conlusion.tex}
    \input{sections/acknowledgements.tex}
    \bibliographystyle{juliacon}
    % \nocite{*}
    \bibliography{Refs}
    \onecolumn
    \input{sections/appendix.tex}
\end{document}