### 2022-01-29 paper/header.tex
Removed backslashes from the paper.yml file.

% **************GENERATED FILE, DO NOT EDIT**************

\title{Causal.jl: A Modeling and Simulation Framework for Causal Models}

\author[1]{Zekeriya Sarı}
\author[1]{Serkan Günel}
\affil[1]{Deparment of Electrical and Electronics Engineering, Dokuz Eylül University, İzmir, TURKEY}

\keywords{System Modeling, Causal Modeling, System Simulation, Dynamical Systems, Julia Programming Language}

\hypersetup{
pdftitle = {Causal.jl: A Modeling and Simulation Framework for Causal Models},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {Zekeriya Sarı, Serkan Günel},
pdfkeywords = {System Modeling, Causal Modeling, System Simulation, Dynamical Systems, Julia Programming Language},
}